Bootstrap learning via modular concept discovery

### 2013-04-22 drafts/ijcai2013/ijcai13.tex
More minor edits.

Genetic programming
(GP)~\cite{DBLP:conf/foga/Koza92}~\cite{DBLP:conf/ices/KozaBAK96} has
tackled the problem of program learning and has explored the notion
that reusable functions are helpful. This work relies, however, on
stochastic local search over the space of programs, and automatically
discovering useful subroutines is seen as a helpful heuristic. More
recently, Liang et al.~\cite{liang10programs} used a stochastic
grammar over combinatory logic to induce shared structure in
multi-task program induction problems. We find this latter
representation compelling and use it here, but we see both this and
genetic programming as tackling the question of how to benefit from
shared structure in situations where local search might be successful
on its own. Such problems rely on domain knowledge -- whether in the
form of a fitness function or a likelihood function -- to provide
local structure to the search space.

But, as mentioned before, the problems we want AI systems to solve
often lack this kind of built in structure. Here we attempt to learn
functional programs to solve those kinds of problems.  \\

@@

We propose the \textbf{E.C. algorithm}, an iterative procedure to solve this
problem. It will be useful in presenting the algorithm to define the
following terminology: a \textbf{frontier} is the finite set of
expressions that the algorithm considers in any one iteration. The
\textbf{frontier size} $N$ is the prespecified number of elements in
the frontier. We will say that a task is \textbf{hit} by the frontier
if there is a expression in the frontier that solves it.

@@

Now suppose that we have a set of expressions $\mathcal{E}$, and we
wish to estimate the maximum likelihood values of the $p_n$'s. If the
choice of primitive at each leaf were not conditioned on the
requesting type, we could just count all occurrences of each primitive
in our expression set, and this would be proportional to the maximum
likelihood estimate. This is the ML estimator for a multinomial
distribution. However, the fact that we condition on the requesting
type makes this a considerably more difficult optimization task. One
of the simplest approximations we can make is to calculate the
frequency with which each primitive combinator appears in
$\mathcal{E}$ when the requesting type is such that it could have
been chosen. This is a straightforward calculation, and
we use it for its convenience, finding that our empirical results
justify its use. Future inquiry will be needed to determine to what
extent a more accurate estimator can improve these results.

@@

Our best-first policy is to expand the next partial solution
with the highest value. If a partial solution has multiple leaves that
are not primitive combinators, we expand the left most one first,
since it is the choice of left combinator that defines the type of its
right combinator in an application pair. That is, the parent node
applies its constraint on valid solutions via the left child, not the
right child. We want to apply constraints as early as possible, so
expanding left child first prunes more partial solutions.

>>>>

Genetic programming
(GP)~\cite{DBLP:conf/foga/Koza92}~\cite{DBLP:conf/ices/KozaBAK96} has
tackled the problem of program learning and has explored the notion
that reusable functions are helpful. This work relies, however, on
stochastic local search over the space of programs, and automatically
discovering useful subroutines is seen as a helpful heuristic. More
recently, Liang et al.~\cite{liang10programs} used a stochastic
grammar over combinatory logic to induce shared structure in
multi-task program induction problems. We find their representation of
programs compelling and use it here, but we see both this and genetic
programming as tackling the question of how to benefit from shared
structure in situations where local search might be successful on its
own. Such problems rely on domain knowledge -- whether in the form of
a fitness function or a likelihood function -- to provide local
structure to the search space.

But, as mentioned before, our goal is to learn to solve problems in
solution spaces whose structure is unknown. Here we attempt to learn
functional programs to solve those kinds of problems.  \\

@@

We propose the \textbf{E.C. algorithm}, an iterative procedure to solve this
problem. It will be useful in presenting the algorithm to define the
following terminology: a \textbf{frontier} is the finite set of
expressions that the algorithm considers in any one iteration. The
\textbf{frontier size} $N$ is the prespecified number of elements in
the frontier. We will say that a task is \textbf{hit} by the frontier
if there is an expression in the frontier that solves it.

@@

Now suppose that we have a set of expressions $\mathcal{E}$, and we
wish to estimate the maximum likelihood values of the $p_n$'s. If the
choice of primitive at each leaf were not conditioned on the
requesting type, we could just count all occurrences of each primitive
in our expression set, and this would be proportional to the maximum
likelihood estimate. This is the ML estimator for a multinomial
distribution. However, the fact that we condition on the requesting
type makes this a considerably more difficult optimization task. A
simple approximation is to calculate the frequency with which each
primitive combinator appears in $\mathcal{E}$ when the requesting type
is such that it could have been chosen. This is a straightforward
calculation, and we use it for its convenience, finding that our
empirical results justify its use. Future inquiry will determine to
what extent a more accurate estimator can improve these results.

@@

Our best-first policy is to expand the next partial solution with the
highest value. If a partial solution has multiple leaves that are not
primitive combinators, we expand the left most one first, since it is
the choice of left combinator that defines the type of its right
combinator in an application pair. That is, the parent node applies
its constraint on valid solutions via the left child, not the right
child. We want to apply constraints as early as possible, and thus we
expand the left child first.
